---
title: "Differential binding of the three PURA HeLa iCLIPs - 2 endo vs oe "
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
  - name: Melina Klostermann
format: 
  html:
    code-fold: true
    code-overflow: scroll
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    
    toc: true
    toc-depth: 3
    toc-location: left
    toc-expand: false
    number-sections: true
    
    theme: sandstone
    fontsize: 11pt
    linestretch: 1.5
        
    fig-format: svg
    cap-location: margin
    crossref:
      fig-title: Fig
    
    embed-resources: true
    link-external-newwindow: true
    smooth-scroll: true
    
    execute:
      echo: true
      warning: false
---



```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=TRUE, tidy.opts=list(width.cutoff=80))

library(ggplot2) 
source("/Users/melinaklostermann/Documents/projects/R_general_functions/theme_thesis.R")
theme_set(theme_thesis())
```




```{r libraries, include=FALSE}

library(dplyr)
library(GenomicFeatures)
library(DESeq2)
library(purrr)


out <- "/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/07_Diff_binding_downstream/"
```

# What was done in this report?

- Overlaps of bound genes and bound regions between the three data sets, on the bases of a union from the binding sites from all three
- MA plots for the differential binding of oe vs endo and flag vs endo
- ECDFs of changes in binding strength, between various sets


```{r}
bs_merge_prefilter <- readRDS("/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/06_Differential_binding/bds_merge_endo_oe_ranges.rds") 
names(bs_merge_prefilter) <- 1:NROW(bs_merge_prefilter)

bs_merge_prefilter <- bs_merge_prefilter |>
  as.data.frame()

bs_merge <- readRDS("/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/06_Differential_binding/merged_bs_diff_endo_oe_res.rds") %>% 
  as.data.frame()

bs_merge2 <- readRDS("/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/06_Differential_binding/merged_bs_diff_flag_oe_res.rds") %>% 
  as.data.frame()

bs_merge3 <- readRDS("/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/06_Differential_binding/merged_bs_diff_endo_flag_res.rds") %>% 
  as.data.frame()

bs_merge4 <- readRDS("/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/06_Differential_binding/merged_bs_diff_oe_2v2_res.rds") %>% 
  as.data.frame()

bs_merge <- bs_merge %>% mutate(
  bs_direction = case_when((bs.padj < 0.01) & (bs.log2FoldChange > 0) ~ "up",
  (bs.padj < 0.01) & (bs.log2FoldChange < 0) ~ "down",
  T ~ "none"),
  bg_direction = case_when((bg.padj < 0.01) & (bg.log2FoldChange > 0) ~ "up",
  (bg.padj < 0.01) & (bg.log2FoldChange < 0) ~ "down",
  T ~ "none")

)


nrow(bs_merge)

# tables
write.csv(as.data.frame(bs_merge), "/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/11_supplementary_tables/S3_Differential_binding_FLAG-PURA_IH-AB_vs_endogenous_PURA.csv")

write.csv(as.data.frame(bs_merge2), "/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/11_supplementary_tables/S4_Differential_binding_FLAG-PURA_FL-AB_vs_endogenous_PURA.csv")

write.csv(as.data.frame(bs_merge4), "/Users/melinaklostermann/Documents/projects/Thesis/Thesis_code/PURA/11_supplementary_tables/S5_Differential_binding_FLAG-PURA_IH-AB_2_vs_2.csv")


```


# Overlaps of binding sites and genes

```{r}
bs_merge_infl <- bs_merge_prefilter %>% tidyr::separate_rows(dataset) %>% split(., .$dataset)
names(bs_merge_infl)

# venn of overlaps
overlaps <- data.frame( endo = bs_merge_prefilter$bsID %in% bs_merge_infl$endo$bsID,
                         oe = bs_merge_prefilter$bsID %in% bs_merge_infl$oe$bsID,
                        flag = bs_merge_prefilter$bsID %in% bs_merge_infl$flag$bsID)
              
 
venn <- eulerr::euler(overlaps)
plot(venn, quantities = TRUE, fontface = 1)

pdf(paste0(out, "venn_bs.pdf"))
plot(venn, quantities = TRUE, fontface = 1)
dev.off()


# venn of overlaps of genes
overlaps2 <- data.frame( endo = unique(bs_merge_prefilter$geneID) %in% bs_merge_infl$endo$geneID,
                         oe = unique(bs_merge_prefilter$geneID) %in% bs_merge_infl$oe$geneID,
                        flag = unique(bs_merge_prefilter$geneID) %in% bs_merge_infl$flag$geneID)
              
 
venn2 <- eulerr::euler(overlaps2)
plot(venn2, quantities = TRUE, fontface = 1)

pdf(paste0(out, "venn_genes.pdf"))
plot(venn2, quantities = TRUE, fontface = 1)
dev.off()

map(bs_merge_infl, ~nrow(.x))
map(bs_merge_infl, ~length(unique(.x$geneID)))

```


# Changes between oe and endo

```{r}

ggplot(bs_merge, aes(x = log10(bs.baseMean), y = bs.log2FoldChange ))+
  ggrastr::rasterise(geom_point(color = "grey", size = 0.25), dpi = 400 )+
   ggrastr::rasterise(geom_point(data = bs_merge %>% subset(., bs.padj < 0.01, ), aes(x = log10(bs.baseMean), y = bs.log2FoldChange ), color = "black", size = 0.25), dpi = 400)+
  theme_thesis()+
  ylim(c(-12,12))

ggsave(paste0(out, "MA_oe_endo.pdf"), width = 6, height = 4, units = "cm")


nrow(bs_merge %>% subset(., bs.padj < 0.01 & bs.log2FoldChange > 0 ))
nrow(bs_merge %>% subset(., bs.padj < 0.01 & bs.log2FoldChange < 0 ))


```



## binding site strength

```{r}

bs_merge_strongest_per_gene <- bs_merge %>% 
  group_by(geneID) %>% 
  arrange(desc(score)) %>%
  dplyr::slice(1)


```



# endo vs flag

```{r}
ggplot(bs_merge3, aes(x = log10(bs.baseMean), y = bs.log2FoldChange ))+
  ggrastr::rasterise(geom_point(color = "grey", size = 0.25), dpi = 400 )+
   ggrastr::rasterise(geom_point(data = bs_merge2 %>% subset(., bs.padj < 0.01, ), aes(x = log10(bs.baseMean), y = bs.log2FoldChange ), color = "black", size = 0.25), dpi = 400)+
  theme_thesis()+
  ylim(c(-12,12))

ggsave(paste0(out, "MA_endo_flag.pdf"), width = 6, height = 4, units = "cm")

nrow(bs_merge3 %>% subset(., bs.padj< 0.01 & bs.log2FoldChange > 0 ))
nrow(bs_merge3 %>% subset(., bs.padj < 0.01 & bs.log2FoldChange < 0 ))


```

# oe vs oe

```{r}
ggplot(bs_merge4, aes(x = log10(bs.baseMean), y = bs.log2FoldChange ))+
  ggrastr::rasterise(geom_point(color = "grey", size = 0.25), dpi = 400 )+
   ggrastr::rasterise(geom_point(data = bs_merge2 %>% subset(., bs.padj < 0.01, ), aes(x = log10(bs.baseMean), y = bs.log2FoldChange ), color = "black", size = 0.25), dpi = 400)+
  theme_thesis()+
  ylim(c(-12,12))

ggsave(paste0(out, "MA_endo_flag.pdf"), width = 6, height = 4, units = "cm")

nrow(bs_merge4 %>% subset(., bs.padj< 0.01 & bs.log2FoldChange > 0 ))
nrow(bs_merge4 %>% subset(., bs.padj < 0.01 & bs.log2FoldChange < 0 ))


```

# Compare all endo-oe flag-oe conditions

```{r}
bs_merge <- left_join(bs_merge, bs_merge2[, c("bsID", "bs.log2FoldChange", "bg.log2FoldChange", "bs.padj")], by = "bsID", suffix = c("", ".flag"))


```





```{r}
bs_merge_strongest_per_gene3 <- bs_merge3 %>% 
  group_by(geneID) %>% 
  arrange(desc(score)) %>%
  dplyr::slice(1)

ggplot(bs_merge, aes( x = bs.log2FoldChange))+
  stat_ecdf(data= bs_merge4, aes(x = bs.log2FoldChange), color = "gray43")+
  stat_ecdf(data= bs_merge3, aes(x = bs.log2FoldChange), color = "#2AB7CA")+
  stat_ecdf(data= bs_merge2, aes(x = bs.log2FoldChange), color = "darkred")+
  stat_ecdf(color = "blue")+
  stat_ecdf(data= bs_merge_strongest_per_gene3, aes(x = bs.log2FoldChange), color = "#2AB7CA")+ # , linetype = "dotted"
  stat_ecdf(data= bs_merge_strongest_per_gene , aes(x = bs.log2FoldChange), color = "blue")+ # , linetype = "dotted"
  coord_cartesian(xlim=c(-5,5))

ggsave(paste0(out, "ecdf_endo_vs_oe_and_flag.pdf"), height = 4, width = 4, units = "cm")


ggplot(bs_merge_strongest_per_gene, aes( x = bs.log2FoldChange))+
  stat_ecdf(data= bs_merge4, aes(x = bs.log2FoldChange), color = "gray43")+
  stat_ecdf(data= bs_merge_strongest_per_gene3, aes(x = bs.log2FoldChange), color = "#2AB7CA")+
  stat_ecdf(color = "blue")+
  coord_cartesian(xlim=c(-5,5))



```



# Session Info
```{r}
sessionInfo()

```
